### 2022-11-24 paper/paper.qmd
Initialize

---
title: "Proof of concept Zstandard custom compression dictionary for FASTA files"
format:
  plos-pdf:
    number-sections: false
    journal:
      # This is an identifier for the target journal: 
      # from https://plos.org/resources/writing-center/ following submission guidelines link, the identifier is the part of the URL after https://journals.plos.org/<id>/s/submission-guidelines
      id: plosbiology
    include-in-header: |
      % Remove comment for double spacing
      % \usepackage{setspace} 
      % \doublespacing
author:
  - name: Michael Persico
    equal-contributor: false
    affiliations:
      - ref: aff1
    notes: |
      "Current Address: Dept/Program/Center, Institution Name, City, State, Country"
    corresponding: false
    email: correspondingauthor@institute.edu
author-notes:
  equal-contributor: These authors contributed equally to this work.
  deceased: Deceased
  group: Membership list can be found in the Acknowledgments sections
affiliations:
  - id: aff1
    name: Department of Biology, Concordia University
    city: Montreal
    state: Quebec
    country: Canada
abstract: |
  ## Background

  Zstandard (Zstd) represents a lossless data compression algorithm that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance. Previous studies paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume. This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements.

  ## Results

  ## Conclusions
bibliography: bibliography.bib 
---

# Introduction

The storage of biological data has represented a significant topic of research, with a number of challenges presented over subsequent generations of technological development, with current challenges discovered as the volume and complexity of data continues to increase[@d2018high;@li2014big].

# Materials and methods

# Results

# Discussion

# Conclusion

# Supporting information

::: {.supp}
## S1 Fig. {#s1-fig}

Bold the title sentence.

Add descriptive text after the title of the item (optional).
:::

::: {.supp}
## S2 Fig. {#s2-fig}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

::: {.supp}
## S1 File. {#s1-file}

Lorem ipsum.

:::

::: {.supp}

## S1 Video. {#s1-video}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

::: {.supp}
## S1 Appendix. {#s1-appendix}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

::: {.supp}
## S1 Table. {#s1-table}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

# Acknowledgments

I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system along with the PLOS template used in this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments; Maximiliano Sandoval and contributors to the Citations app  for managing the paper's bibliography, and all other persons that had indirectly assisted through their programs and research.

### 2022-11-24 paper/paper.tex
Initialize

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%

\PassOptionsToPackage{table}{xcolor}

\documentclass[
  10pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% marvosym package for additional characters
\usepackage{marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% Using natbib instead
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Proof of concept Zstandard custom compression dictionary for FASTA files},
  pdfauthor={Michael Persico},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Proof of concept Zstandard custom compression dictionary
for FASTA
files} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Michael Persico\textsuperscript{1}
\\
\bigskip
\textbf{1} Department of Biology, Concordia
University, Montreal, Quebec, Canada, 
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments
sections

% Use the asterisk to denote corresponding authorship and provide email address in note below.

\end{flushleft}

\section*{Abstract}
\hypertarget{background}{%
\subsection{Background}\label{background}}

Zstandard (Zstd) represents a lossless data compression algorithm that
is highly configurable and is aimed at coupling high compression ratios
with fast compression/decompression performance. Previous studies paired
specific Zstd configurations with various file formats in bioinformatics
to reduce total data volume. This paper presents a ``training mode''
pipeline, written in the Julia programming language, wherein a custom
compression dictionary is generated from a sample FASTA set in order to
explore further compression improvements.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

\linenumbers\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The storage of biological data has represented a significant topic of
research, with a number of challenges presented over subsequent
generations of technological development, with current challenges
discovered as the volume and complexity of data continues to
increase\citep{d2018high, li2014big}.

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

\paragraph*{S1 Fig.}
\label{s1-fig}
{\textbf{Bold the title sentence.}} Add descriptive text after the title
of the item (optional).

\paragraph*{S2 Fig.}
\label{s2-fig}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 File.}
\label{s1-file}
{\textbf{Lorem ipsum.}}

\paragraph*{S1 Video.}
\label{s1-video}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{s1-appendix}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 Table.}
\label{s1-table}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

I would like to express my sincere gratitude to Professor David Walsh
for his advice that helped shape the research as it progressed; My
friends and family that supported me throughout my studies; The Julia
community for their support before and throughout the writing of this
paper; The Quarto developers for producing the Quarto publishing system
along with the PLOS template used in this paper; Luca Di Maio and
contributors to the Distrobox tool for the ease of setting up the
containerized development environments; Maximiliano Sandoval and
contributors to the Citations app for managing the paper's bibliography,
and all other persons that had indirectly assisted through their
programs and research.

\nolinenumbers
  \bibliography{bibliography.bib}

\end{document}

### 2022-11-24 src/FinalProjectBIOL480.jl

@@ -1,0 +1,8 @@
+module FinalProjectBIOL480
+
+using PyCall
+
+zstd = pyimport_conda("pyzstd", "pyzstd=0.15.3")
+
+
+end # module FinalProjectBIOL480

### 2022-11-24 dev/GenerateArtifacts.jl
Add DataFrames dep, Artifacts.toml

@@ -1,0 +1,9 @@
+#=
+Script to generate Julia artifacts, which represent
+fixed, immutable containers for data, in this instance
+any dataset utilized for creating Zstd dictionaries. All 
+artifacts are located in the artifacts/ directory, and 
+their entries are listed in the Artifacts.toml file.
+Docs: https://pkgdocs.julialang.org/v1/artifacts/
+=#
+

### 2022-11-24 src/Persico2022.jl

@@ -1,8 +1,8 @@
-module FinalProjectBIOL480
+module Persico2022
 
 using PyCall
 
 zstd = pyimport_conda("pyzstd", "pyzstd=0.15.3")
-
+println(zstd.zstd_version_info)
 
 end # module FinalProjectBIOL480

### 2022-11-24 paper/paper.qmd
Add CodecLz4

@@
 ---
-title: "Proof of concept Zstandard custom compression dictionary for FASTA files"
+title: "Proof of concept Zstandard user-provided dictionary compression for FASTA files"
 format:
@@
 
-  Zstandard (Zstd) represents a lossless data compression algorithm that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
-Previous studies paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume.
-This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements.
+  Zstandard (Zstd) represents a lossless data compression mechanism that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
+Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume.
+This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 compressors.
 
@@
 # Materials and methods
+
+A list of direct and indirect dependencies can be found in the repository's Manifest.toml file.
 
@@
 
-I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system along with the PLOS template used in this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments; Maximiliano Sandoval and contributors to the Citations app  for managing the paper's bibliography, and all other persons that had indirectly assisted through their programs and research.
+I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.

### 2022-11-24 paper/paper.tex
Add CodecLz4

@@
 \hypersetup{
-  pdftitle={Proof of concept Zstandard custom compression dictionary for FASTA files},
+  pdftitle={Proof of concept Zstandard user-provided dictionary compression for FASTA files},
   pdfauthor={Michael Persico},
@@
 {\Large
-\textbf\newline{Proof of concept Zstandard custom compression dictionary
-for FASTA
+\textbf\newline{Proof of concept Zstandard user-provided dictionary
+compression for FASTA
 files} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
@@
 
-Zstandard (Zstd) represents a lossless data compression algorithm that
+Zstandard (Zstd) represents a lossless data compression mechanism that
 is highly configurable and is aimed at coupling high compression ratios
-with fast compression/decompression performance. Previous studies paired
-specific Zstd configurations with various file formats in bioinformatics
-to reduce total data volume. This paper presents a ``training mode''
-pipeline, written in the Julia programming language, wherein a custom
-compression dictionary is generated from a sample FASTA set in order to
-explore further compression improvements.
+with fast compression/decompression performance. Previous studies have
+paired specific Zstd configurations with various file formats in
+bioinformatics to reduce total data volume. This paper presents a
+``training mode'' pipeline, written in the Julia programming language,
+wherein a custom compression dictionary is generated from a sample FASTA
+set in order to explore further compression improvements and compare
+them to the compression performance of Xz, Zlib, Bzip2, and Lz4
+compressors.
 
@@
 \section{Materials and methods}\label{materials-and-methods}}
+
+A list of direct and indirect dependencies can be found in the
+repository's Manifest.toml file.
 
@@
 community for their support before and throughout the writing of this
-paper; The Quarto developers for producing the Quarto publishing system
-along with the PLOS template used in this paper; Luca Di Maio and
-contributors to the Distrobox tool for the ease of setting up the
-containerized development environments; Maximiliano Sandoval and
-contributors to the Citations app for managing the paper's bibliography,
-and all other persons that had indirectly assisted through their
-programs and research.
+paper; The Quarto developers for producing the Quarto publishing
+system\citep{Allaire_Quarto_2022} along with the PLOS template used for
+this paper; Luca Di Maio and contributors to the Distrobox tool for the
+ease of setting up the containerized development
+environments\citep{maio}; Maximiliano Sandoval and contributors to the
+Citations app for managing the paper's bibliography\citep{sandoval}, and
+all other persons that had indirectly assisted through their programs
+and research.
 

### 2022-11-25 paper/paper.qmd
Change to mkitti CodecZstd branch + expand Introduction

@@
 
-  Zstandard (Zstd) represents a lossless data compression mechanism that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
+  Zstandard (Zstd) represents a universal, lossless data compression mechanism that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
 Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume.
-This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 compressors.
+This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 universal compressors.
 
@@
 
-The storage of biological data has represented a significant topic of research, with a number of challenges presented over subsequent generations of technological development, with current challenges discovered as the volume and complexity of data continues to increase[@d2018high;@li2014big].
+The explosion of biological data has represented a significant topic of research, with a number of challenges presented over subsequent generations of technological development in regards to the management of  the increasing volume and complexity of data[@d2018high;@li2014big].
+In response, emerging trends in data management have lead to the development of novel, scalable methods for the efficient transmission and storage of large amounts of data[@sais2022intelligent].
+With a potentially exponential quantity of files, datasets, and other data resources to be handled, data compression represents a method for reducing overall resource size by encoding the original data into a compact form, thus helping to ease storage requirements [@jayasankar2021survey].
+Research into data compression in the context of biological data began to pick up near the turn of the 21st century as universal compression algorithms at the time were not considered ideal for compressing DNA or RNA sequence data well, which led to the introduction of purpose-built algorithms that addressed the unique attributes of genomics data[@grumbach1994new].
+At the same time, new file formats were introduced, either text-based or binary-based, for more accurate structuring and representation of biological data, complementing new software tools[@lipman1985rapid;@mills2014common].
+
+The FASTA file format is a legacy of the original FASTA program from 1985 for finding sequence similarities with a query sequence[@lipman1985rapid].
+Each file can possess multiple sequences, each paired with a description line distinguished with a ">" symbol followed by arbitary text, usually a name and summary description, on the same line.
+It is a commonly supported file format in bioinformatics and has been the target for optimized data compressors in competition with one another.
+The DELIMINATE lossless algorithm was first proposed in 2012, wherein header and sequence data are separated into DELIM-1 and DELIM-2 variants and a two-phase process is pursued involving delta encoding, progressive elimination of nucleotide characters, and 7-Zip archiving[@mohammed2012deliminate].
+The claims of better compression/decompression performance of FASTA files were soon rivaled by the introduction of the MFCompress tool, again separating headers and sequence data but instead relying on probabilistic models to encode the data[@pinho2014mfcompress], which was countered by the Nucleotide Archival Format, a novel file format which is noteworthy in this context for the inclusion a Zstandard compression step[@kryukov2019nucleotide].
 
@@
 
-A list of direct and indirect dependencies can be found in the repository's Manifest.toml file.
+A list of all direct and indirect dependencies can be found in the repository's Manifest.toml file.
 

### 2022-11-25 paper/paper.tex
Change to mkitti CodecZstd branch + expand Introduction

@@
 
-Zstandard (Zstd) represents a lossless data compression mechanism that
-is highly configurable and is aimed at coupling high compression ratios
-with fast compression/decompression performance. Previous studies have
-paired specific Zstd configurations with various file formats in
-bioinformatics to reduce total data volume. This paper presents a
-``training mode'' pipeline, written in the Julia programming language,
-wherein a custom compression dictionary is generated from a sample FASTA
-set in order to explore further compression improvements and compare
-them to the compression performance of Xz, Zlib, Bzip2, and Lz4
-compressors.
+Zstandard (Zstd) represents a universal, lossless data compression
+mechanism that is highly configurable and is aimed at coupling high
+compression ratios with fast compression/decompression performance.
+Previous studies have paired specific Zstd configurations with various
+file formats in bioinformatics to reduce total data volume. This paper
+presents a ``training mode'' pipeline, written in the Julia programming
+language, wherein a custom compression dictionary is generated from a
+sample FASTA set in order to explore further compression improvements
+and compare them to the compression performance of Xz, Zlib, Bzip2, and
+Lz4 universal compressors.
 
@@
 
-The storage of biological data has represented a significant topic of
+The explosion of biological data has represented a significant topic of
 research, with a number of challenges presented over subsequent
-generations of technological development, with current challenges
-discovered as the volume and complexity of data continues to
-increase\citep{d2018high, li2014big}.
+generations of technological development in regards to the management of
+the increasing volume and complexity of
+data\citep{d2018high, li2014big}. In response, emerging trends in data
+management have lead to the development of novel, scalable methods for
+the efficient transmission and storage of large amounts of
+data\citep{sais2022intelligent}. With a potentially exponential quantity
+of files, datasets, and other data resources to be handled, data
+compression represents a method for reducing overall resource size by
+encoding the original data into a compact form, thus helping to ease
+storage requirements \citep{jayasankar2021survey}. Research into data
+compression in the context of biological data began to pick up near the
+turn of the 21st century as universal compression algorithms at the time
+were not considered ideal for compressing DNA or RNA sequence data well,
+which led to the introduction of purpose-built algorithms that addressed
+the unique attributes of genomics data\citep{grumbach1994new}. At the
+same time, new file formats were introduced, either text-based or
+binary-based, for more accurate structuring and representation of
+biological data, complementing new software
+tools\citep{lipman1985rapid, mills2014common}.
+
+The FASTA file format is a legacy of the original FASTA program from
+1985 for finding sequence similarities with a query
+sequence\citep{lipman1985rapid}. Each file can possess multiple
+sequences, each paired with a description line distinguished with a
+``\textgreater{}'' symbol followed by arbitary text, usually a name and
+summary description, on the same line. It is a commonly supported file
+format in bioinformatics and has been the target for optimized data
+compressors in competition with one another. The DELIMINATE lossless
+algorithm was first proposed in 2012, wherein header and sequence data
+are separated into DELIM-1 and DELIM-2 variants and a two-phase process
+is pursued involving delta encoding, progressive elimination of
+nucleotide characters, and 7-Zip
+archiving\citep{mohammed2012deliminate}. The claims of better
+compression/decompression performance of FASTA files were soon rivaled
+by the introduction of the MFCompress tool, again separating headers and
+sequence data but instead relying on probabilistic models to encode the
+data\citep{pinho2014mfcompress}, which was countered by the Nucleotide
+Archival Format, a novel file format which is noteworthy in this context
+for the inclusion a Zstandard compression
+step\citep{kryukov2019nucleotide}.
 
@@
 
-A list of direct and indirect dependencies can be found in the
+A list of all direct and indirect dependencies can be found in the
 repository's Manifest.toml file.

### 2022-11-25 src/Persico2022.jl
Remove PyCall + CodecZstd dict tree branch citation + gitignore for all dotfiles

@@ -1,8 +1,5 @@
 module Persico2022
 
-using PyCall
-
-zstd = pyimport_conda("pyzstd", "pyzstd=0.15.3")
-println(zstd.zstd_version_info)
+println("Hello World!")
 
 end # module FinalProjectBIOL480

### 2022-11-26 dev/GenerateArtifacts.jl
Add Artifacts

@@ -1,9 +1,0 @@
-#=
-Script to generate Julia artifacts, which represent
-fixed, immutable containers for data, in this instance
-any dataset utilized for creating Zstd dictionaries. All 
-artifacts are located in the artifacts/ directory, and 
-their entries are listed in the Artifacts.toml file.
-Docs: https://pkgdocs.julialang.org/v1/artifacts/
-=#
-

### 2022-11-26 src/GenerateArtifacts.jl

@@ -1,0 +1,18 @@
+"""
+Generate Julia artifacts, which represent fixed, immutable containers for data, in 
+this instance any dataset utilized for creating Zstd dictionaries. Major advantages include 
+better reproducibility, and automatic tarball extraction. Each entry is listed in the 
+Artifacts.toml file. Docs: https://pkgdocs.julialang.org/v1/artifacts/
+"""
+
+using Pkg.Artifacts
+using ArtifactUtils
+
+function artifact_generate(name::String, url::String)
+    add_artifact!(Artifacts.find_artifacts_toml(@__DIR__), name , url, lazy = true)
+    ensure_artifact_installed(name, Artifacts.find_artifacts_toml(@__DIR__))
+    #archive dataset in he datasets/ directory in case it is removed from its original location or file server goes down
+    archive_artifact(artifact_hash(name, find_artifacts_toml(@__DIR__)), joinpath(dirname(@__DIR__), "datasets/$(name)_archive.tar.gz"))
+end
+
+artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")

### 2022-11-26 src/Persico2022.jl

@@ -1,5 +1,3 @@
 module Persico2022
 
-println("Hello World!")
-
-end # module FinalProjectBIOL480
+end 

### 2022-11-26 src/GenerateArtifacts.jl
Artifact refactoring

@@ -1,7 +1,7 @@
 """
 Generate Julia artifacts, which represent fixed, immutable containers for data, in 
 this instance any dataset utilized for creating Zstd dictionaries. Major advantages include 
-better reproducibility, and automatic tarball extraction. Each entry is listed in the 
+better reproducibility, and automatic tarball unpacking. Each entry is listed in the 
 Artifacts.toml file. Docs: https://pkgdocs.julialang.org/v1/artifacts/
 """
 
@@ -9,10 +9,9 @@ using Pkg.Artifacts
 using ArtifactUtils
 
 function artifact_generate(name::String, url::String)
-    add_artifact!(Artifacts.find_artifacts_toml(@__DIR__), name , url, lazy = true)
+    add_artifact!(Artifacts.find_artifacts_toml(@__DIR__), name , url, lazy = false, force = true)
     ensure_artifact_installed(name, Artifacts.find_artifacts_toml(@__DIR__))
-    #archive dataset in he datasets/ directory in case it is removed from its original location or file server goes down
-    archive_artifact(artifact_hash(name, find_artifacts_toml(@__DIR__)), joinpath(dirname(@__DIR__), "datasets/$(name)_archive.tar.gz"))
 end
 
+# Generate the E.coli *.fna (DNA sequence) NCBI dataset 
 artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")

### 2022-11-26 src/Persico2022.jl

@@ -1,3 +1,5 @@
 module Persico2022
 
+include("GenerateArtifacts.jl")
+
 end 

### 2022-11-28 paper/paper.qmd
Finish intro

@@
 ---
-title: "Proof of concept Zstandard user-provided dictionary compression for FASTA files"
+title: "Exploring Zstandard user-provided dictionary compression for FASTA files"
 format:
@@
 
-  Zstandard (Zstd) represents a universal, lossless data compression mechanism that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
+  Zstandard (Zstd) represents a universal, lossless data compression standard and implementation that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
 Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume.
@@
 With a potentially exponential quantity of files, datasets, and other data resources to be handled, data compression represents a method for reducing overall resource size by encoding the original data into a compact form, thus helping to ease storage requirements [@jayasankar2021survey].
-Research into data compression in the context of biological data began to pick up near the turn of the 21st century as universal compression algorithms at the time were not considered ideal for compressing DNA or RNA sequence data well, which led to the introduction of purpose-built algorithms that addressed the unique attributes of genomics data[@grumbach1994new].
+Research into data compression in the context of biological data began to pick up near the turn of the 21st century as universal compression algorithms at the time were not considered ideal for compressing DNA or RNA sequence data well, which led to the introduction of purpose-built algorithms that addressed the unique peculiarities of genomic data[@grumbach1994new].
 At the same time, new file formats were introduced, either text-based or binary-based, for more accurate structuring and representation of biological data, complementing new software tools[@lipman1985rapid;@mills2014common].
 
-The FASTA file format is a legacy of the original FASTA program from 1985 for finding sequence similarities with a query sequence[@lipman1985rapid].
-Each file can possess multiple sequences, each paired with a description line distinguished with a ">" symbol followed by arbitary text, usually a name and summary description, on the same line.
-It is a commonly supported file format in bioinformatics and has been the target for optimized data compressors in competition with one another.
+The FASTA file format is a legacy of the original FASTA program for finding sequence similarities with a query sequence[@lipman1985rapid].
+Each file can possess multiple sequences, each paired with a description line distinguished by a ">" symbol followed by arbitrary text, usually a name and/or summary description, on the same line.
+It is a commonly supported file format in bioinformatics and has been the target for optimized data compressors with competing claims for performance.
 The DELIMINATE lossless algorithm was first proposed in 2012, wherein header and sequence data are separated into DELIM-1 and DELIM-2 variants and a two-phase process is pursued involving delta encoding, progressive elimination of nucleotide characters, and 7-Zip archiving[@mohammed2012deliminate].
-The claims of better compression/decompression performance of FASTA files were soon rivaled by the introduction of the MFCompress tool, again separating headers and sequence data but instead relying on probabilistic models to encode the data[@pinho2014mfcompress], which was countered by the Nucleotide Archival Format, a novel file format which is noteworthy in this context for the inclusion a Zstandard compression step[@kryukov2019nucleotide].
+The claims of better compression/decompression performance of FASTA files were soon rivaled by the introduction of the MFCompress tool, again separating headers and sequence data but instead relying on probabilistic models to encode the data[@pinho2014mfcompress], which was then countered by the Nucleotide Archival Format, a novel file format noteworthy in this context for the inclusion of a Zstandard compression step[@kryukov2019nucleotide].
+
+IETF RFC 8878, introduced by engineers at Facebook, defines Zstandard as a lossless data compression/decompression format[@collet2021rfc8478].
+It is often abbreviated as "Zstd", though such can also refer to Facebook's own implementation of the algorithm written mostly in C[@facebook].
+Content is sliced and packaged into "frames" that are independent of one another defined as either compressed data Zstandard frames or Skippable frames containing custom user metadata[@collet2021rfc8478].
+Zstd's backbone is the use of Finite State Entropy and Huffman entropy encoding schemes that replace data with coded forms independently of the medium[@ezhilarasan_thambidurai_praveena_srinivasan_sumathi_2007;@LU20171], with the former compressing all symbols, though header information is first encoded by the latter[@collet2021rfc8478].
+Zstd, for small data compression improvements, also functions as a dictionary coder, meaning that although the algorithm is universal in the sense of being applicable to a number of both text-based and binary-based data files, it can be optimized for compacting characteristic data by "training" Zstd with a collection of sample files to build a set of common patterns that allow for substitutions when compressing/decompressing, allowing for further  gains for similar data[@8316293].
+
+Common bioinformatics file formats often include a set structure with the express purpose of representing specific kinds of biological data composed of repeating elements, as is the case with FASTA files with either nucleotide or amino acid sequences.
+In this work is described a pipeline for building Zstd dictionaries via FASTA datasets along with a comparison of Zstd compression/decompression performance with that of several alternative lossless compressors for select datasets.
 

### 2022-11-28 paper/paper.tex
Finish intro

@@
 \hypersetup{
-  pdftitle={Proof of concept Zstandard user-provided dictionary compression for FASTA files},
+  pdftitle={Exploring Zstandard user-provided dictionary compression for FASTA files},
   pdfauthor={Michael Persico},
@@
 {\Large
-\textbf\newline{Proof of concept Zstandard user-provided dictionary
-compression for FASTA
+\textbf\newline{Exploring Zstandard user-provided dictionary compression
+for FASTA
 files} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
@@
 Zstandard (Zstd) represents a universal, lossless data compression
-mechanism that is highly configurable and is aimed at coupling high
-compression ratios with fast compression/decompression performance.
-Previous studies have paired specific Zstd configurations with various
-file formats in bioinformatics to reduce total data volume. This paper
-presents a ``training mode'' pipeline, written in the Julia programming
-language, wherein a custom compression dictionary is generated from a
-sample FASTA set in order to explore further compression improvements
-and compare them to the compression performance of Xz, Zlib, Bzip2, and
-Lz4 universal compressors.
+standard and implementation that is highly configurable and is aimed at
+coupling high compression ratios with fast compression/decompression
+performance. Previous studies have paired specific Zstd configurations
+with various file formats in bioinformatics to reduce total data volume.
+This paper presents a ``training mode'' pipeline, written in the Julia
+programming language, wherein a custom compression dictionary is
+generated from a sample FASTA set in order to explore further
+compression improvements and compare them to the compression performance
+of Xz, Zlib, Bzip2, and Lz4 universal compressors.
 
@@
 which led to the introduction of purpose-built algorithms that addressed
-the unique attributes of genomics data\citep{grumbach1994new}. At the
+the unique peculiarities of genomic data\citep{grumbach1994new}. At the
 same time, new file formats were introduced, either text-based or
@@
 
-The FASTA file format is a legacy of the original FASTA program from
-1985 for finding sequence similarities with a query
+The FASTA file format is a legacy of the original FASTA program for
+finding sequence similarities with a query
 sequence\citep{lipman1985rapid}. Each file can possess multiple
-sequences, each paired with a description line distinguished with a
-``\textgreater{}'' symbol followed by arbitary text, usually a name and
-summary description, on the same line. It is a commonly supported file
-format in bioinformatics and has been the target for optimized data
-compressors in competition with one another. The DELIMINATE lossless
-algorithm was first proposed in 2012, wherein header and sequence data
-are separated into DELIM-1 and DELIM-2 variants and a two-phase process
-is pursued involving delta encoding, progressive elimination of
-nucleotide characters, and 7-Zip
+sequences, each paired with a description line distinguished by a
+``\textgreater{}'' symbol followed by arbitrary text, usually a name
+and/or summary description, on the same line. It is a commonly supported
+file format in bioinformatics and has been the target for optimized data
+compressors with competing claims for performance. The DELIMINATE
+lossless algorithm was first proposed in 2012, wherein header and
+sequence data are separated into DELIM-1 and DELIM-2 variants and a
+two-phase process is pursued involving delta encoding, progressive
+elimination of nucleotide characters, and 7-Zip
 archiving\citep{mohammed2012deliminate}. The claims of better
@@
 sequence data but instead relying on probabilistic models to encode the
-data\citep{pinho2014mfcompress}, which was countered by the Nucleotide
-Archival Format, a novel file format which is noteworthy in this context
-for the inclusion a Zstandard compression
+data\citep{pinho2014mfcompress}, which was then countered by the
+Nucleotide Archival Format, a novel file format noteworthy in this
+context for the inclusion of a Zstandard compression
 step\citep{kryukov2019nucleotide}.
+
+IETF RFC 8878, introduced by engineers at Facebook, defines Zstandard as
+a lossless data compression/decompression
+format\citep{collet2021rfc8478}. It is often abbreviated as ``Zstd'',
+though such can also refer to Facebook's own implementation of the
+algorithm written mostly in C\citep{facebook}. Content is sliced and
+packaged into ``frames'' that are independent of one another defined as
+either compressed data Zstandard frames or Skippable frames containing
+custom user metadata\citep{collet2021rfc8478}. Zstd's backbone is the
+use of Finite State Entropy and Huffman entropy encoding schemes that
+replace data with coded forms independently of the
+medium\citep{ezhilarasan_thambidurai_praveena_srinivasan_sumathi_2007, LU20171},
+with the former compressing all symbols, though header information is
+first encoded by the latter\citep{collet2021rfc8478}. Zstd, for small
+data compression improvements, also functions as a dictionary coder,
+meaning that although the algorithm is universal in the sense of being
+applicable to a number of both text-based and binary-based data files,
+it can be optimized for compacting characteristic data by ``training''
+Zstd with a collection of sample files to build a set of common patterns
+that allow for substitutions when compressing/decompressing, allowing
+for further gains for similar data\citep{8316293}.
+
+Common bioinformatics file formats often include a set structure with
+the express purpose of representing specific kinds of biological data
+composed of repeating elements, as is the case with FASTA files with
+either nucleotide or amino acid sequences. In this work is described a
+pipeline for building Zstd dictionaries via FASTA datasets along with a
+comparison of Zstd compression/decompression performance with that of
+several alternative lossless compressors for select datasets.
 

### 2022-12-05 paper/paper.qmd
Start Methods, Discussion sections

@@
 Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume.
-This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 universal compressors.
+This paper presents a primitive "training mode" model, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 universal compressors.
 
@@
 
-A list of all direct and indirect dependencies can be found in the repository's Manifest.toml file.
+For the model implementation, Julia was chosen as there is package support for working with sequence data as well as for trivial transcoding of compressed or decompressed data via the TranscodingStreams package wherein compression algorithm implementations are loaded as codec I/O streams[@sato].
+A list of all direct dependencies can be found in the repository's Project.toml file.
+The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into an artifact, or life-cycled datastore.
 
@@
 
-# Conclusion
+The provided FASTA training dataset was retrieved from a set of \textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome shotgun run[@ncbi].
+Sampling a single organism may allow for additional compression/decompression improvements as the sequence data could possess repeating patterns unique to that individual's or species' genome that are recognized and added to the Zstd dictionary during training.
+The FASTA files themselves may bear similarity with one another in the same set, such as each file being of similar length.
+Simultaneously, the Zstd dictionary may not achieve comparable compression/decompression performance when attempted on more dissimilar sequence data.
+Portability of Zstd dictionaries, in the sense of being universally applicable across all sequence data, could be weighed against generating dedicated dictionaries for specific datasets depending on the storage requirements.
+Zstd parameters were kept default, though they can be manually assigned, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].
 

### 2022-12-05 paper/paper.tex
Start Methods, Discussion sections

@@
 with various file formats in bioinformatics to reduce total data volume.
-This paper presents a ``training mode'' pipeline, written in the Julia
-programming language, wherein a custom compression dictionary is
+This paper presents a primitive ``training mode'' model, written in the
+Julia programming language, wherein a custom compression dictionary is
 generated from a sample FASTA set in order to explore further
@@
 
-A list of all direct and indirect dependencies can be found in the
-repository's Manifest.toml file.
+For the model implementation, Julia was chosen as there is package
+support for working with sequence data as well as for trivial
+transcoding of compressed or decompressed data via the
+TranscodingStreams package wherein compression algorithm implementations
+are loaded as codec I/O streams\citep{sato}. A list of all direct
+dependencies can be found in the repository's Project.toml file. The
+Julia implementation operates by first unloading a tarball containing
+the set of FASTA training files into an artifact, or life-cycled
+datastore.
 
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+The provided FASTA training dataset was retrieved from a set of
+\textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome
+shotgun run\citep{ncbi}. Sampling a single organism may allow for
+additional compression/decompression improvements as the sequence data
+could possess repeating patterns unique to that individual's or species'
+genome that are recognized and added to the Zstd dictionary during
+training. The FASTA files themselves may bear similarity with one
+another in the same set, such as each file being of similar length.
+Simultaneously, the Zstd dictionary may not achieve comparable
+compression/decompression performance when attempted on more dissimilar
+sequence data. Portability of Zstd dictionaries, in the sense of being
+universally applicable across all sequence data, could be weighed
+against generating dedicated dictionaries for specific datasets
+depending on the storage requirements. Zstd parameters were kept
+default, though they can be manually assigned, including advanced
+compression options like compression job size or selecting from 22
+predefined compression levels\citep{facebook}.
 

### 2022-12-05 src/GenerateArtifacts.jl

@@ -12,6 +12,3 @@ function artifact_generate(name::String, url::String)
     add_artifact!(Artifacts.find_artifacts_toml(@__DIR__), name , url, lazy = false, force = true)
     ensure_artifact_installed(name, Artifacts.find_artifacts_toml(@__DIR__))
 end
-
-# Generate the E.coli *.fna (DNA sequence) NCBI dataset 
-artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")

### 2022-12-05 src/Persico2022.jl

@@ -2,4 +2,8 @@ module Persico2022
 
 include("GenerateArtifacts.jl")
 
+# Generate the E.coli *.fna (DNA sequence) NCBI dataset 
+artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")
+
+
 end 

### 2022-12-05 src/Pipeline.jl

@@ -1,0 +1,10 @@
+#=
+=#
+
+using TranscodingStreams
+using CodecBzip2
+using CodecLz4
+using CodecXz
+using CodecZlib
+using CodecZstd
+ 

### 2022-12-05 src/Stats.jl

@@ -1,0 +1,4 @@
+#=
+Generate statistics for the resulting compression/decompression data 
+(tables, plots).
+=#

### 2022-12-06 src/Pipeline.jl
Backbone for tests

@@ -1,5 +1,5 @@
-#=
-=#
+"""
+"""
 
 using TranscodingStreams
 using CodecBzip2

### 2022-12-06 src/Plots.jl

@@ -1,0 +1,21 @@
+"""
+Generate plots for visualization of the compression trends and other information
+"""
+
+using VegaLite
+using VegaDatasets
+using StatsBase
+
+include("Sequences.jl")
+
+df = compression_tests(:AA, 100, 100000)
+
+println()
+
+p = df |>
+@vlplot(
+    mark={:boxplot, extent="min-max"},
+    x="type:o",
+    y={:size, axis={title="population"}}
+)
+save("figure.png", p)

### 2022-12-06 src/Sequences.jl

@@ -1,0 +1,89 @@
+"""
+Set of functions that allow for generating FASTA formatted data for running compression
+tests using Xz, Bzip2, Zlib, and Zstd compressors
+"""
+
+using FASTX
+using BioSequences
+using DataFrames
+using TranscodingStreams
+using CodecXz
+using CodecBzip2
+using CodecZlib
+using CodecZstd
+using StatsBase
+
+
+"""
+Generate a Random FASTA file with a random DNA sequence length from range 1 to length
+"""
+function FASTA_DNA_generate(min_length::Int, max_length::Int)
+    # Generate random DNA sequence
+    seq = BioSequences.randdnaseq(rand(min_length:max_length))
+    # Convert DNA sequence to FASTA format
+    rec = FASTA.Record("seq", seq)
+    return rec.data
+end
+
+"""
+Generate a Random FASTA file with a random RNA sequence length from range 1 to length
+"""
+function FASTA_RNA_generate(min_length::Int, max_length::Int)
+    # Generate random DNA sequence
+    seq = BioSequences.randrnaseq(rand(min_length:max_length))
+    # Convert DNA sequence to FASTA format
+    rec = FASTA.Record("seq", seq)
+    return rec.data
+end
+
+"""
+Generate a Random FASTA file with a random amino acid sequence length from range 1 to length
+"""
+function FASTA_AA_generate(min_length::Int, max_length::Int)
+    # Generate random DNA sequence
+    seq = BioSequences.randaaseq(rand(min_length:max_length))
+    # Convert DNA sequence to FASTA format
+    rec = FASTA.Record("seq", seq)
+    return rec.data
+end
+
+"""
+Generate compression size information from randomly generated FASTA data
+"""
+function compression_tests(type::Symbol, 
+    files::Int, 
+    min_length::Int, 
+    max_length::Int,
+    )
+
+    size_df = DataFrame(
+        type = Symbol[],
+        size = Int64[],
+    )
+    for i = 1:files
+        
+        # Can generate either DNA, RNA, or amino acid sequence data
+        if type == :DNA 
+            data = FASTA_DNA_generate(min_length, max_length) 
+        elseif type ==:RNA
+            data = FASTA_RNA_generate(min_length, max_length)
+        elseif type ==:AA
+            data = FASTA_AA_generate(min_length, max_length)
+        else
+            return error("Invalid sequence type!")
+        end
+
[…]

### 2022-12-07 paper/paper.qmd
Add workflow plot, more README changes

@@
 
-For the model implementation, Julia was chosen as there is package support for working with sequence data as well as for trivial transcoding of compressed or decompressed data via the TranscodingStreams package wherein compression algorithm implementations are loaded as codec I/O streams[@sato].
+All resources for the paper are included in the public Github repository at github.com/M-PERSIC/Persico2022.git.
+
+The model contains two pipelines, one for Zstd custom dictionary generation and another for generating transcoded compression data using Bzip2, Xz, zLib, and Zstd compressors, with a general overview in @fig-model.
+Julia was chosen as there is package support for working with sequence data as well as for trivial transcoding of compressed or decompressed data via TranscodingStreams.jl wherein compression algorithm implementations are loaded as codecs[@sato].
 A list of all direct dependencies can be found in the repository's Project.toml file.
-The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into an artifact, or life-cycled datastore.
+The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into a Julia artifact, or life-cycled datastore.
+The files are plugged into Zstd's Dictionary API to generate a custom dictionary specific to that dataset.
+Random FASTA data is generated using BioSequences.jl[@biosequences] and FASTX.jl[@fastx] APIs which is then transcoded into compressed formats via the CodecBzip2.jl[@codecbzip2], CodecXz.jl[@codecxz], CodecZlib.jl[@codeczlib], and CodecZstd.jl[@codeczstd] codecs for comparative analysis.
+Both dictionary-based and default Zstd compression are applied to determine if the former desmontrates a significant reduction in data size.
+
+![Overview of the Julia implementation for the compression model.
+**A**, Custom dictionary generation via an Julia artifact of FASTA files plugged into the Zstd Dictionary API.
+**B**, Transcoding randomly generated FASTA formatted data using TranscodingStreams.jl[@sato] compression codecs.
+](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/figures/Compression_Pipeline_Model.png){#fig-model}
 
@@
 Portability of Zstd dictionaries, in the sense of being universally applicable across all sequence data, could be weighed against generating dedicated dictionaries for specific datasets depending on the storage requirements.
-Zstd parameters were kept default, though they can be manually assigned, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].
+Zstd parameters were kept default, though they can be manually altered, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].
 

### 2022-12-07 paper/paper.tex
Add workflow plot, more README changes

@@
 
-For the model implementation, Julia was chosen as there is package
-support for working with sequence data as well as for trivial
-transcoding of compressed or decompressed data via the
-TranscodingStreams package wherein compression algorithm implementations
-are loaded as codec I/O streams\citep{sato}. A list of all direct
-dependencies can be found in the repository's Project.toml file. The
-Julia implementation operates by first unloading a tarball containing
-the set of FASTA training files into an artifact, or life-cycled
-datastore.
+All resources for the paper are included in the public Github repository
+at github.com/M-PERSIC/Persico2022.git.
+
+The model contains two pipelines, one for Zstd custom dictionary
+generation and another for generating transcoded compression data using
+Bzip2, Xz, zLib, and Zstd compressors, with a general overview in
+Fig~\ref{fig-model}. Julia was chosen as there is package support for
+working with sequence data as well as for trivial transcoding of
+compressed or decompressed data via TranscodingStreams.jl wherein
+compression algorithm implementations are loaded as codecs\citep{sato}.
+A list of all direct dependencies can be found in the repository's
+Project.toml file. The Julia implementation operates by first unloading
+a tarball containing the set of FASTA training files into a Julia
+artifact, or life-cycled datastore. The files are plugged into Zstd's
+Dictionary API to generate a custom dictionary specific to that dataset.
+Random FASTA data is generated using BioSequences.jl\citep{biosequences}
+and FASTX.jl\citep{fastx} APIs which is then transcoded into compressed
+formats via the CodecBzip2.jl\citep{codecbzip2},
+CodecXz.jl\citep{codecxz}, CodecZlib.jl\citep{codeczlib}, and
+CodecZstd.jl\citep{codeczstd} codecs for comparative analysis. Both
+dictionary-based and default Zstd compression are applied to determine
+if the former desmontrates a significant reduction in data size.
+
+\begin{figure}
+
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/figures/Compression_Pipeline_Model.png}
+
+}
+
+\caption{\label{fig-model}Overview of the Julia implementation for the
+compression model. \textbf{A}, Custom dictionary generation via an Julia
+artifact of FASTA files plugged into the Zstd Dictionary API.
+\textbf{B}, Transcoding randomly generated FASTA formatted data using
+TranscodingStreams.jl\citep{sato} compression codecs.}
+
+\end{figure}
 
@@
 depending on the storage requirements. Zstd parameters were kept
-default, though they can be manually assigned, including advanced
+default, though they can be manually altered, including advanced
 compression options like compression job size or selecting from 22

### 2022-12-07 src/Pipeline.jl

@@ -1,10 +1,0 @@
-"""
-"""
-
-using TranscodingStreams
-using CodecBzip2
-using CodecLz4
-using CodecXz
-using CodecZlib
-using CodecZstd
- 

### 2022-12-07 src/Sequences.jl

@@ -15,7 +15,7 @@ using StatsBase
 
 
 """
-Generate a Random FASTA file with a random DNA sequence length from range 1 to length
+Generate FASTA formatted record with a random DNA sequence with a random length between two bounds
 """
 function FASTA_DNA_generate(min_length::Int, max_length::Int)
     # Generate random DNA sequence
@@ -26,23 +26,23 @@ function FASTA_DNA_generate(min_length::Int, max_length::Int)
 end
 
 """
-Generate a Random FASTA file with a random RNA sequence length from range 1 to length
+Generate FASTA formatted record with a random RNA sequence with a random length between two bounds
 """
 function FASTA_RNA_generate(min_length::Int, max_length::Int)
-    # Generate random DNA sequence
+    # Generate random RNA sequence
     seq = BioSequences.randrnaseq(rand(min_length:max_length))
-    # Convert DNA sequence to FASTA format
+    # Convert RNA sequence to FASTA format
     rec = FASTA.Record("seq", seq)
     return rec.data
 end
 
 """
-Generate a Random FASTA file with a random amino acid sequence length from range 1 to length
+Generate FASTA formatted record with a random amino acid sequence with a random length between two bounds
 """
 function FASTA_AA_generate(min_length::Int, max_length::Int)
-    # Generate random DNA sequence
+    # Generate random AA sequence
     seq = BioSequences.randaaseq(rand(min_length:max_length))
-    # Convert DNA sequence to FASTA format
+    # Convert AA sequence to FASTA format
     rec = FASTA.Record("seq", seq)
     return rec.data
 end

### 2022-12-09 paper/paper.qmd
Major changes, added Zstd dictionary functions

@@
 author:
-  - name: Michael Persico
+  - name: Michael A. Persico
     equal-contributor: false
@@
 
-All resources for the paper are included in the public Github repository at github.com/M-PERSIC/Persico2022.git.
+All resources for the paper are included in the public Github repository at [github.com/M-PERSIC/Persico2022.git](https://www.github.com/M-PERSIC/Persico2022.git).
 
@@
 The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into a Julia artifact, or life-cycled datastore.
-The files are plugged into Zstd's Dictionary API to generate a custom dictionary specific to that dataset.
+The files are plugged into Zstd's training mode, via the Zstd command line interface --train option, to generate a custom dictionary specific to that dataset.
+Subsequent dictionary manipulation in Julia is based on an experimental CodecZstd.jl branch that includes functions that wrap the Zstd Dictionary API[@kittisopikul].
 Random FASTA data is generated using BioSequences.jl[@biosequences] and FASTX.jl[@fastx] APIs which is then transcoded into compressed formats via the CodecBzip2.jl[@codecbzip2], CodecXz.jl[@codecxz], CodecZlib.jl[@codeczlib], and CodecZstd.jl[@codeczstd] codecs for comparative analysis.
@@
 Zstd parameters were kept default, though they can be manually altered, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].
+
+Zstd's multifarious design warrants further research in how tuning the implementation by altering available options (block size, compression level,...) that might influence compression performance.
+The development of novel data toolsets have utilized Zstd in unique ways.
+Aside from the Nucleotide Archival Format, FASTAFS was recently introduced as a filesystem in userspace (FUSE) toolkit wherein FASTA files are pooled into a virtualization abstraction layer that archives the set of data synced with its metadata, generating a high-level interface for accessing and storing FASTA data[@fastafs].
+The zstd-seekable compression library is an embedded component that keeps the FASTAFS archive compressed for storage reduction purposes, and it represents an alternative compressed data format that deviates from Zstandard, albeit in a compatible manner, that allows for partial decompression of subranges of data[@facebook].
+This is achieved via independently compressed/decompressed Zstd frames, which FASTAFS exploits for fast random access to FASTA files by decompressing only data of interest.
+Further areas of research for general lossless data compression in the context of biological data could include exploring other properties like compression/decompression speed performance, meaning how fast data transcoding occurs, or piloting proof-of-concept algorithms that attempt novel strategies.
 
@@
 
-I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.
+I would like to express my sincere gratitude to Professor David Walsh for teaching the bioinformatics course; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; Mark Kittisopikul for developing the CodecZstd.jl #dictionary_and_parameters branch for Zstd dictionary functions; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.
+
+# Conflict of interest
+
+The author declares no conflict of interest.

### 2022-12-09 paper/paper.tex
Major changes, added Zstd dictionary functions

@@
   pdftitle={Exploring Zstandard user-provided dictionary compression for FASTA files},
-  pdfauthor={Michael Persico},
+  pdfauthor={Michael A. Persico},
   hidelinks,
@@
 % Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-Michael Persico\textsuperscript{1}
+Michael A. Persico\textsuperscript{1}
 \\
@@
 All resources for the paper are included in the public Github repository
-at github.com/M-PERSIC/Persico2022.git.
+at
+\href{https://www.github.com/M-PERSIC/Persico2022.git}{github.com/M-PERSIC/Persico2022.git}.
 
@@
 artifact, or life-cycled datastore. The files are plugged into Zstd's
-Dictionary API to generate a custom dictionary specific to that dataset.
-Random FASTA data is generated using BioSequences.jl\citep{biosequences}
-and FASTX.jl\citep{fastx} APIs which is then transcoded into compressed
-formats via the CodecBzip2.jl\citep{codecbzip2},
-CodecXz.jl\citep{codecxz}, CodecZlib.jl\citep{codeczlib}, and
-CodecZstd.jl\citep{codeczstd} codecs for comparative analysis. Both
-dictionary-based and default Zstd compression are applied to determine
-if the former desmontrates a significant reduction in data size.
+training mode, via the Zstd command line interface --train option, to
+generate a custom dictionary specific to that dataset. Subsequent
+dictionary manipulation in Julia is based on an experimental
+CodecZstd.jl branch that includes functions that wrap the Zstd
+Dictionary API\citep{kittisopikul}. Random FASTA data is generated using
+BioSequences.jl\citep{biosequences} and FASTX.jl\citep{fastx} APIs which
+is then transcoded into compressed formats via the
+CodecBzip2.jl\citep{codecbzip2}, CodecXz.jl\citep{codecxz},
+CodecZlib.jl\citep{codeczlib}, and CodecZstd.jl\citep{codeczstd} codecs
+for comparative analysis. Both dictionary-based and default Zstd
+compression are applied to determine if the former desmontrates a
+significant reduction in data size.
 
@@
 predefined compression levels\citep{facebook}.
+
+Zstd's multifarious design warrants further research in how tuning the
+implementation by altering available options (block size, compression
+level,\ldots) that might influence compression performance. The
+development of novel data toolsets have utilized Zstd in unique ways.
+Aside from the Nucleotide Archival Format, FASTAFS was recently
+introduced as a filesystem in userspace (FUSE) toolkit wherein FASTA
+files are pooled into a virtualization abstraction layer that archives
+the set of data synced with its metadata, generating a high-level
+interface for accessing and storing FASTA data\citep{fastafs}. The
+zstd-seekable compression library is an embedded component that keeps
+the FASTAFS archive compressed for storage reduction purposes, and it
+represents an alternative compressed data format that deviates from
+Zstandard, albeit in a compatible manner, that allows for partial
+decompression of subranges of data\citep{facebook}. This is achieved via
+independently compressed/decompressed Zstd frames, which FASTAFS
+exploits for fast random access to FASTA files by decompressing only
+data of interest. Further areas of research for general lossless data
+compression in the context of biological data could include exploring
+other properties like compression/decompression speed performance,
+meaning how fast data transcoding occurs, or piloting proof-of-concept
+algorithms that attempt novel strategies.
 
@@
 I would like to express my sincere gratitude to Professor David Walsh
-for his advice that helped shape the research as it progressed; My
-friends and family that supported me throughout my studies; The Julia
-community for their support before and throughout the writing of this
-paper; The Quarto developers for producing the Quarto publishing
+for teaching the bioinformatics course; My friends and family that
+supported me throughout my studies; The Julia community for their
+support before and throughout the writing of this paper; Mark
+Kittisopikul for developing the CodecZstd.jl
+\#dictionary\_and\_parameters branch for Zstd dictionary functions; The
+Quarto developers for producing the Quarto publishing
 system\citep{Allaire_Quarto_2022} along with the PLOS template used for
@@
 
+\hypertarget{conflict-of-interest}{%
+\section{Conflict of interest}\label{conflict-of-interest}}
+
+The author declares no conflict of interest.
+
 \nolinenumbers

### 2022-12-09 src/Dictionary.jl

@@ -1,0 +1,29 @@
+"""
+Set of functions for creating the custom Zstd compression dictionary.
+"""
+
+using Pkg.Artifacts
+using CodecZstd
+
+"""
+Create the Zstd dictionary by running the Zstd CLI implementation.
+"""
+function create_dictionary(artifact::String)
+    run(Cmd(`sudo apt update`))
+    run(Cmd(`sudo apt install -y zstd`))
+    run(Cmd(`zstd --train -r "$artifact"`))
+    return (CodecZstd.Dictionary(read("dictionary")))
+end
+
+"""
+Generate Zstd Compressor Stream object
+"""
+create_cstream() = CodecZstd.CStream()
+
+"""
+Load the Zstd Dictionary for CodecZstd usage.
+"""
+function load_dictionary(cstream::CodecZstd.CStream, dict::CodecZstd.Dictionary)
+    CodecZstd.loadDictionary(cstream, dict)
+end
+

### 2022-12-09 src/Persico2022.jl

@@ -3,7 +3,13 @@ module Persico2022
 include("GenerateArtifacts.jl")
 
 # Generate the E.coli *.fna (DNA sequence) NCBI dataset 
-artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")
+#artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")
 
+include("Sequences.jl")
+df = compression_tests(:DNA, 100, 100000, 150000)
+
+include("Plots.jl")
+
+compression_performance_plot_generate(df, "In the name of all that is glorious", "meuw", "muow", "assets/figures/meuw.png")
 
 end 

### 2022-12-09 src/Plots.jl

@@ -1,5 +1,5 @@
 """
-Generate plots for visualization of the compression trends and other information
+Generate plots for visualization of the compression trends and other trends
 """
 
 using VegaLite
@@ -8,14 +8,57 @@ using StatsBase
 
 include("Sequences.jl")
 
-df = compression_tests(:AA, 100, 100000)
+df = compression_tests(:DNA, 100, 100000, 150000)
 
-println()
-
-p = df |>
+df |>
 @vlplot(
+    width=240,
+    height=135,
     mark={:boxplot, extent="min-max"},
-    x="type:o",
-    y={:size, axis={title="population"}}
+    title="muow",
+    x={"type:o", axis={title="maow"}},
+    y={:size, axis={title="meuw"}},
 )
-save("figure.png", p)
+
+"""
+Generate whisker plot for a visual demonstration of the range of sizes for compressed FASTA data
+"""
+function compression_performance_plot_generate(df::DataFrame, 
+    title::String,
+    xtitle::String,
+    ytitle::String,
+    filepath::String)
+whiskerplot = df |>
+@vlplot(
+    width=240,
+    height=135,
+    mark={:boxplot, extent="min-max"},
+    title=title,
+    x={"type:o", axis={title=xtitle}},
+    y={:size, axis={title=ytitle}},
+)
+save(filepath, whiskerplot)
+end
+
+"""
+Generate bar plot for a visual demonstration of the total size of all files following compression
+"""
+function total_filesize_plot_generate(df::DataFrame,
+    title::String,
+    xtitle::String,
+    ytitle::String,
+    filepath::String
+    )
+barplot = df |>
+@vlplot(
+    width=240,
+    height=135,
+    mark={:boxplot, extent="min-max"},
+    title=title,
+    x={"type:o", axis={title=xtitle}},
+    y={:size, aggregate="sum", axis={title=title}},
+)
+save(filepath, barplot)
+end
+
+function 

### 2022-12-09 src/Sequences.jl

@@ -1,6 +1,6 @@
 """
 Set of functions that allow for generating FASTA formatted data for running compression
-tests using Xz, Bzip2, Zlib, and Zstd compressors
+tests using Xz, Bzip2, Zlib, and Zstd compressors.
 """
 
 using FASTX
@@ -11,7 +11,7 @@ using CodecXz
 using CodecBzip2
 using CodecZlib
 using CodecZstd
-using StatsBase
+using Pkg.Artifacts
 
 
 """
@@ -50,18 +50,14 @@ end
 """
 Generate compression size information from randomly generated FASTA data
 """
-function compression_tests(type::Symbol, 
-    files::Int, 
-    min_length::Int, 
-    max_length::Int,
-    )
+function compression_tests(type::Symbol, files::Int, min_length::Int, max_length::Int, artifact::String)
 
     size_df = DataFrame(
         type = Symbol[],
         size = Int64[],
     )
+
     for i = 1:files
-        
         # Can generate either DNA, RNA, or amino acid sequence data
         if type == :DNA 
             data = FASTA_DNA_generate(min_length, max_length) 
@@ -82,8 +78,19 @@ function compression_tests(type::Symbol,
         push!(size_df, [:Xz, data_xz_compressed])
         data_zlib_compressed = sizeof(transcode(ZlibCompressor, data))
         push!(size_df, [:Zlib, data_zlib_compressed])
-        data_zstd_compressed = sizeof(transcode(ZstdCompressor, data))
-        push!(size_df, [:Zstd_default, data_zstd_compressed])
+        data_zstd_default_compressed = sizeof(transcode(ZstdCompressor, data))
+        push!(size_df, [:Zstd_default, data_zstd_default_compressed])
+
+        # Train, load custom Zstd dictionary
+        dict = create_dictionary(artifact"ecoli_MS_200")
+        cstream = create_cstream()
+        load_dictionary(cstream, dict)
+        data_zstd_dict_compressed = sizeof(transcode(ZstdCompressor, data))
+        push!(size_df, [:Zstd_dict, data_zstd_dict_compressed])
+
+        # Clear custom dictionary from CodecZstd for rerun if desired
+        rm("dictionary")
+        CodecZstd.reset!(cstream, 0)
     end
     return size_df
 end

### 2022-12-10 paper/paper.qmd
Final code

@@
 The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into a Julia artifact, or life-cycled datastore.
-The files are plugged into Zstd's training mode, via the Zstd command line interface --train option, to generate a custom dictionary specific to that dataset.
-Subsequent dictionary manipulation in Julia is based on an experimental CodecZstd.jl branch that includes functions that wrap the Zstd Dictionary API[@kittisopikul].
+The files are plugged into Zstd's training mode, via the Zstd command line interface --train option, to generate a custom dictionary specific to that dataset that is then plugged again into the CLI for data compression.
 Random FASTA data is generated using BioSequences.jl[@biosequences] and FASTX.jl[@fastx] APIs which is then transcoded into compressed formats via the CodecBzip2.jl[@codecbzip2], CodecXz.jl[@codecxz], CodecZlib.jl[@codeczlib], and CodecZstd.jl[@codeczstd] codecs for comparative analysis.
-Both dictionary-based and default Zstd compression are applied to determine if the former desmontrates a significant reduction in data size.
+Both dictionary-based and default Zstd compression are applied to determine if the former results in a significant reduction of overall data size.
 
@@
 **A**, Custom dictionary generation via an Julia artifact of FASTA files plugged into the Zstd Dictionary API.
-**B**, Transcoding randomly generated FASTA formatted data using TranscodingStreams.jl[@sato] compression codecs.
-](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/figures/Compression_Pipeline_Model.png){#fig-model}
+**B**, Transcoding randomly generated FASTA formatted data using TranscodingStreams.jl[@sato] compression codecs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/diagrams/Compression_Model_Overview.png){#fig-model}
 
 # Results
+
+![Average FASTA formatted data size (bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression.
+*Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation.
+Randomized DNA sequences within the range of 100 000 to 150 000 base pairs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png){#fig-performance}
 
@@
 
-# Supporting information
+# Supplementary information
 
@@
 
-::: {.supp}
-## S1 File. {#s1-file}
-
-Lorem ipsum.
-
-:::
-
-::: {.supp}
-
-## S1 Video. {#s1-video}
-
-Lorem ipsum.
-
-Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-:::
-
-::: {.supp}
-## S1 Appendix. {#s1-appendix}
-
-Lorem ipsum.
-
-Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-:::
-
-::: {.supp}
-## S1 Table. {#s1-table}
-
-Lorem ipsum.
-
-Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-:::
-
 # Acknowledgments

### 2022-12-10 paper/paper.tex
Final code

@@
 training mode, via the Zstd command line interface --train option, to
-generate a custom dictionary specific to that dataset. Subsequent
-dictionary manipulation in Julia is based on an experimental
-CodecZstd.jl branch that includes functions that wrap the Zstd
-Dictionary API\citep{kittisopikul}. Random FASTA data is generated using
-BioSequences.jl\citep{biosequences} and FASTX.jl\citep{fastx} APIs which
-is then transcoded into compressed formats via the
-CodecBzip2.jl\citep{codecbzip2}, CodecXz.jl\citep{codecxz},
-CodecZlib.jl\citep{codeczlib}, and CodecZstd.jl\citep{codeczstd} codecs
-for comparative analysis. Both dictionary-based and default Zstd
-compression are applied to determine if the former desmontrates a
-significant reduction in data size.
+generate a custom dictionary specific to that dataset that is then
+plugged again into the CLI for data compression. Random FASTA data is
+generated using BioSequences.jl\citep{biosequences} and
+FASTX.jl\citep{fastx} APIs which is then transcoded into compressed
+formats via the CodecBzip2.jl\citep{codecbzip2},
+CodecXz.jl\citep{codecxz}, CodecZlib.jl\citep{codeczlib}, and
+CodecZstd.jl\citep{codeczstd} codecs for comparative analysis. Both
+dictionary-based and default Zstd compression are applied to determine
+if the former results in a significant reduction of overall data size.
 
@@
 
-{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/figures/Compression_Pipeline_Model.png}
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/diagrams/Compression_Model_Overview.png}
 
@@
 \section{Results}\label{results}}
+
+\begin{figure}
+
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png}
+
+}
+
+\caption{\label{fig-performance}Average FASTA formatted data size
+(bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based
+Zstd compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
+artifact for Zstd dictionary generation. Randomized DNA sequences within
+the range of 100 000 to 150 000 base pairs.}
+
+\end{figure}
 
@@
 
-\hypertarget{supporting-information}{%
-\section{Supporting information}\label{supporting-information}}
+\hypertarget{supplementary-information}{%
+\section{Supplementary information}\label{supplementary-information}}
 
@@
 \label{s2-fig}
-{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
-gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
-ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
-consectetur pellentesque.
-
-\paragraph*{S1 File.}
-\label{s1-file}
-{\textbf{Lorem ipsum.}}
-
-\paragraph*{S1 Video.}
-\label{s1-video}
-{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
-gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
-ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
-consectetur pellentesque.
-
-\paragraph*{S1 Appendix.}
-\label{s1-appendix}
-{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
-gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
-ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
-consectetur pellentesque.
-
-\paragraph*{S1 Table.}
-\label{s1-table}
 {\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices

### 2022-12-10 src/Data.jl

@@ -1,0 +1,39 @@
+"""
+Set of functions for randomly generating FASTA-formatted data containing either DNA, RNA, or amino acid sequences
+"""
+
+using FASTX
+using BioSequences
+
+"""
+Generate FASTA formatted record with a random DNA sequence with a random length between two bounds
+"""
+function FASTA_DNA_generate(min_length::Int, max_length::Int)
+    # Generate random DNA sequence
+    seq = BioSequences.randdnaseq(rand(min_length:max_length))
+    # Convert DNA sequence to FASTA format
+    rec = FASTA.Record("seq", seq)
+    return rec.data
+end
+
+"""
+Generate FASTA formatted record with a random RNA sequence with a random length between two bounds
+"""
+function FASTA_RNA_generate(min_length::Int, max_length::Int)
+    # Generate random RNA sequence
+    seq = BioSequences.randrnaseq(rand(min_length:max_length))
+    # Convert RNA sequence to FASTA format
+    rec = FASTA.Record("seq", seq)
+    return rec.data
+end
+
+"""
+Generate FASTA formatted record with a random amino acid sequence with a random length between two bounds
+"""
+function FASTA_AA_generate(min_length::Int, max_length::Int)
+    # Generate random AA sequence
+    seq = BioSequences.randaaseq(rand(min_length:max_length))
+    # Convert AA sequence to FASTA format
+    rec = FASTA.Record("seq", seq)
+    return rec.data
+end

### 2022-12-10 src/Dictionary.jl

@@ -1,29 +1,0 @@
-"""
-Set of functions for creating the custom Zstd compression dictionary.
-"""
-
-using Pkg.Artifacts
-using CodecZstd
-
-"""
-Create the Zstd dictionary by running the Zstd CLI implementation.
-"""
-function create_dictionary(artifact::String)
-    run(Cmd(`sudo apt update`))
-    run(Cmd(`sudo apt install -y zstd`))
-    run(Cmd(`zstd --train -r "$artifact"`))
-    return (CodecZstd.Dictionary(read("dictionary")))
-end
-
-"""
-Generate Zstd Compressor Stream object
-"""
-create_cstream() = CodecZstd.CStream()
-
-"""
-Load the Zstd Dictionary for CodecZstd usage.
-"""
-function load_dictionary(cstream::CodecZstd.CStream, dict::CodecZstd.Dictionary)
-    CodecZstd.loadDictionary(cstream, dict)
-end
-

### 2022-12-10 src/Persico2022.jl

@@ -1,15 +1,59 @@
 module Persico2022
 
-include("GenerateArtifacts.jl")
+"""
+Zstd custom dictionary compression model. A compression dictionary is generated with an E.coli dataset and then compression performance is compared across Bzip2, Xz, Zlib, Zstd (default parameters), and Zstd via the custom dictionary.
+"""
 
-# Generate the E.coli *.fna (DNA sequence) NCBI dataset 
-#artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")
+using DataFrames
 
-include("Sequences.jl")
-df = compression_tests(:DNA, 100, 100000, 150000)
+# Generate the E.coli *.fna (DNA sequence) NCBI dataset artifact
+include("Dataset.jl")
+artifact_generate("ecoli_MS_200", "https://ftp.ncbi.nih.gov/genomes/HUMAN_MICROBIOM/Bacteria/Escherichia_coli_MS_200_1_uid47275/NZ_ADUC00000000.scaffold.fna.tgz")
 
+
+# Run the compression tests on randomized FASTA data containing DNA sequences, input the results in a DataFrame object
+# The Zstd dictionary will be generated inside the compression_tests function
+include("Tests.jl")
+df = compression_tests(:DNA, 100, 100000, 150000, artifact"ecoli_MS_200")
+
+# Generate plots of the data compression trends.
 include("Plots.jl")
 
-compression_performance_plot_generate(df, "In the name of all that is glorious", "meuw", "muow", "assets/figures/meuw.png")
+# Main result plot comparing compression performance of random DNA FASTA formatted data
+compression_performance_whiskerplot_generate(df, 
+"Data Compression Size Distributions (E.Coli dataset[24] training dataset, DNA length = 100 000 - 150 000)",
+"Data compressor type", 
+"Data size in bytes",
+"/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/PlotOne.png")
+
+# Plot the TukeyHSD statistical results in a table 
+include("Stats.jl")
+tukey_test_generate(df, "var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/data/results.csv")
+tukey_results = csv_to_df("/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/data/results.csv")
[…]

### 2022-12-10 src/Plots.jl

@@ -2,63 +2,49 @@
 Generate plots for visualization of the compression trends and other trends
 """
 
-using VegaLite
-using VegaDatasets
-using StatsBase
-
-include("Sequences.jl")
-
-df = compression_tests(:DNA, 100, 100000, 150000)
-
-df |>
-@vlplot(
-    width=240,
-    height=135,
-    mark={:boxplot, extent="min-max"},
-    title="muow",
-    x={"type:o", axis={title="maow"}},
-    y={:size, axis={title="meuw"}},
-)
+using DataFrames
+using PlotlyJS
 
 """
 Generate whisker plot for a visual demonstration of the range of sizes for compressed FASTA data
 """
-function compression_performance_plot_generate(df::DataFrame, 
-    title::String,
-    xtitle::String,
-    ytitle::String,
-    filepath::String)
-whiskerplot = df |>
-@vlplot(
-    width=240,
-    height=135,
-    mark={:boxplot, extent="min-max"},
-    title=title,
-    x={"type:o", axis={title=xtitle}},
-    y={:size, axis={title=ytitle}},
-)
-save(filepath, whiskerplot)
+function compression_performance_whiskerplot_generate(df::DataFrame, title::String, x_title::String, y_title::String, filepath::String)
+    uncompressed_data = df[(df.type .== "Uncompressed"), :size]
+    uncompressed_trace = box(y = uncompressed_data, boxpoints = "all", kind = "box", name = "Uncompressed", marker_color = "rgb(105, 105, 105)", notched = true)
+    
+    bzip2_data = df[(df.type .== "Bzip2"), :size]
+    bzip2_trace = box(y = bzip2_data, boxpoints = "all", kind = "box", name = "Bzip2", marker_color = "rgb(255, 0, 255)", notched = true)
+    
+    xz_data = df[(df.type .== "Xz"), :size]
+    xz_trace = box(y = xz_data, boxpoints = "all", kind = "box", name = "Xz", marker_color = "rgb(0, 255, 255)", notched = true)
+    
+    zlib_data = df[(df.type .== "Zlib"), :size]
+    zlib_trace = box(y = zlib_data, boxpoints = "all", kind = "box", name = "Zlib", marker_color = "rgb(0, 255, 0)", notched = true)
+    
+    zstd_default_data = df[(df.type .== "Zstd_default"), :size]
[…]

### 2022-12-10 src/Sequences.jl

@@ -1,96 +1,0 @@
-"""
-Set of functions that allow for generating FASTA formatted data for running compression
-tests using Xz, Bzip2, Zlib, and Zstd compressors.
-"""
-
-using FASTX
-using BioSequences
-using DataFrames
-using TranscodingStreams
-using CodecXz
-using CodecBzip2
-using CodecZlib
-using CodecZstd
-using Pkg.Artifacts
-
-
-"""
-Generate FASTA formatted record with a random DNA sequence with a random length between two bounds
-"""
-function FASTA_DNA_generate(min_length::Int, max_length::Int)
-    # Generate random DNA sequence
-    seq = BioSequences.randdnaseq(rand(min_length:max_length))
-    # Convert DNA sequence to FASTA format
-    rec = FASTA.Record("seq", seq)
-    return rec.data
-end
-
-"""
-Generate FASTA formatted record with a random RNA sequence with a random length between two bounds
-"""
-function FASTA_RNA_generate(min_length::Int, max_length::Int)
-    # Generate random RNA sequence
-    seq = BioSequences.randrnaseq(rand(min_length:max_length))
-    # Convert RNA sequence to FASTA format
-    rec = FASTA.Record("seq", seq)
-    return rec.data
-end
-
-"""
-Generate FASTA formatted record with a random amino acid sequence with a random length between two bounds
-"""
-function FASTA_AA_generate(min_length::Int, max_length::Int)
-    # Generate random AA sequence
-    seq = BioSequences.randaaseq(rand(min_length:max_length))
-    # Convert AA sequence to FASTA format
-    rec = FASTA.Record("seq", seq)
-    return rec.data
-end
-
-"""
-Generate compression size information from randomly generated FASTA data
-"""
-function compression_tests(type::Symbol, files::Int, min_length::Int, max_length::Int, artifact::String)
-
-    size_df = DataFrame(
-        type = Symbol[],
-        size = Int64[],
-    )
-
-    for i = 1:files
-        # Can generate either DNA, RNA, or amino acid sequence data
-        if type == :DNA 
-            data = FASTA_DNA_generate(min_length, max_length) 
-        elseif type ==:RNA
-            data = FASTA_RNA_generate(min_length, max_length)
-        elseif type ==:AA
-            data = FASTA_AA_generate(min_length, max_length)
-        else
-            return error("Invalid sequence type!")
-        end
-
[…]

### 2022-12-10 src/Stats.jl

@@ -1,4 +1,37 @@
-#=
-Generate statistics for the resulting compression/decompression data 
-(tables, plots).
-=#
+"""
+Set of functions for aiding with generating a Tukey-test for statistical comparison of compression performance
+"""
+
+using RCall
+using DataFrames
+using CSV
+
+include("Tests.jl")
+
+"""
+Use The TukeyHSD R package to perform a Tukey's HSD (honestly significant difference) test
+"""
+function tukey_test_generate(df::DataFrame, filepath::String)
+    df_types = robject(df[:,"type"])
+    df_sizes = robject(df[:, "size"])
+    @rput df_types
+    @rput df_sizes
+    R"""
+    df <- do.call(rbind.data.frame, Map('c', df_types, df_sizes))
+    colnames(df) <- c("type", "size")
+    model <- aov(size ~ type, data=df)
+    tukey_test <- TukeyHSD(model, conf.level=.95)
+    library(broom)
+    data <- tidy(tukey_test)
+    write.csv(data, file="assets/data/results.csv")
+    """
+end
+
+"""
+Helper function to convert TukeyHSD test results, saved as a .csv file, into a DataFrame in order to plot as a table
+"""
+csv_to_df(filepath::String) = CSV.read(filepath, DataFrame)
+
+
+
+

### 2022-12-10 src/Tests.jl

@@ -1,0 +1,77 @@
+"""
+All-in-one function for running compression tests between the different Bzip2, Xz, Zlib, Zstd (default parameters), and Zstd via the custom dictionary compressors. 
+"""
+
+using DataFrames
+using Pkg.Artifacts
+using TranscodingStreams
+using CodecXz
+using CodecBzip2
+using CodecZlib
+using CodecZstd
+
+include("Data.jl")
+
+"""
+Create the Zstd dictionary by running the Zstd CLI implementation
+"""
+function create_dictionary(artifact::String)
+    run(Cmd(`sudo apt update`))
+    run(Cmd(`sudo apt install -y zstd`))
+    run(Cmd(`zstd --train -r "$artifact"`))
+end
+
+
+"""
+Generate compression size information from randomly generated FASTA data
+"""
+function compression_tests(type::Symbol, files::Int, min_length::Int, max_length::Int, artifact::String)
+
+    # DataFrame object to hold all of the compression data
+    size_df = DataFrame(
+        type = String[],
+        size = Int64[],
+    )
+
+    # Generate Zstd custom dictionary
+    create_dictionary(artifact"ecoli_MS_200")
+
+    for i = 1:files
+        # Can generate either DNA, RNA, or amino acid sequence data
+        if type == :DNA 
+            data = FASTA_DNA_generate(min_length, max_length) 
+        elseif type ==:RNA
+            data = FASTA_RNA_generate(min_length, max_length)
+        elseif type ==:AA
+            data = FASTA_AA_generate(min_length, max_length)
+        else
+            return error("Invalid sequence type!")
+        end
+
+        # Get uncompressed/compressed sizes (kilobytes) of the data and record in dataframe
+        data_uncompressed_size = sizeof(data)
+        push!(size_df, ["Uncompressed", data_uncompressed_size])
+        data_bzip2_compressed = sizeof(transcode(Bzip2Compressor, data))
+        push!(size_df, ["Bzip2", data_bzip2_compressed])
+        data_xz_compressed = sizeof(transcode(XzCompressor, data))
+        push!(size_df, ["Xz", data_xz_compressed])
+        data_zlib_compressed = sizeof(transcode(ZlibCompressor, data))
+        push!(size_df, ["Zlib", data_zlib_compressed])
+        data_zstd_default_compressed = sizeof(transcode(ZstdCompressor, data))
+        push!(size_df, ["Zstd_default", data_zstd_default_compressed])
+
[…]

### 2022-12-10 paper/Persico2022.qmd
Final paper, final code

--- 2022-12-10 paper/paper.qmd
@@
 abstract: |
-  ## Background
-
   Zstandard (Zstd) represents a universal, lossless data compression standard and implementation that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance.
@@
 This paper presents a primitive "training mode" model, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 universal compressors.
+Zstd dictionary-based compression did not yield significant compression gains (P < 0.05 in many compressor contrast scenarios using Tukey's HSD(honestly significant difference) test) and had performed worse than alternative compressor types.
+As an *E.coli* training dataset was used to generate a dictionary to compress randomized data, and with the abundance of options available for manually tuning Zstd's parameters, further compression improvements may be possible.
 
-  ## Results
-
-  ## Conclusions
 bibliography: bibliography.bib 
@@
 
+\newpage
+
 # Results
 
-![Average FASTA formatted data size (bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression.
-*Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation.
-Randomized DNA sequences within the range of 100 000 to 150 000 base pairs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png){#fig-performance}
+![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression.
+*Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation[23].
+Randomized DNA sequences within the range between 100 000 to 150 000 base pairs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png){#fig-performance}
+
+![Tukey's HSD test for determining statistical significance between the compression performance, based on compressed data size, of Bzip2, Xz, Zlib, Zstd with default parameters, and Zstd with a custom dictionary based on the E.coli training dataset[23].](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Tukeytable.png){#fig-tukey}
+
+\newpage
+
+The Tukey's HSD test paints a picture of the statistical significance of the means of the compressed set of FASTA formatted data between each compressor type.
+As expected, uncompressed data size and compressed data size differ significantly, with no compressed FASTA file above more than 60 kilobytes based on @fig-performance.
+With dictionary-based compression applied, Zstd performs similarly to Zstd compression with default parameters (P > 0.05) and, based on the trends of the boxplot and the Tukey's test results, in almost every instance aside from default Zstd compression, dictionary-based compression performed worse than the competing compressor types.
 
@@
 
+As stated in the Results section, Zstd dictionary-based compression results were surprising in the sense that, at least with the configuration (particular *E.coli* dataset, specific number of randomized FASTA files, completely randomized DNA, RNA, or amino acid sequences), it performed either on par, as was the case with default Zstd compression, or worse than the alternative compressors.
 The provided FASTA training dataset was retrieved from a set of \textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome shotgun run[@ncbi].
@@
 The FASTA files themselves may bear similarity with one another in the same set, such as each file being of similar length.
-Simultaneously, the Zstd dictionary may not achieve comparable compression/decompression performance when attempted on more dissimilar sequence data.
+Simultaneously, the Zstd dictionary may not achieve comparable compression/decompression performance when attempted on more dissimilar sequence data, which might explain the observed results as randomized sequences were generated for compression performance analysis.
 Portability of Zstd dictionaries, in the sense of being universally applicable across all sequence data, could be weighed against generating dedicated dictionaries for specific datasets depending on the storage requirements.
-Zstd parameters were kept default, though they can be manually altered, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].
+All other Zstd parameters were kept default, though they can be manually altered, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].
 
-Zstd's multifarious design warrants further research in how tuning the implementation by altering available options (block size, compression level,...) that might influence compression performance.
+Zstd's multifarious design warrants further research in how tuning the implementation by altering available options (block size, compression level,...) might yield more significant compression performance.
 The development of novel data toolsets have utilized Zstd in unique ways.
@@
 This is achieved via independently compressed/decompressed Zstd frames, which FASTAFS exploits for fast random access to FASTA files by decompressing only data of interest.
-Further areas of research for general lossless data compression in the context of biological data could include exploring other properties like compression/decompression speed performance, meaning how fast data transcoding occurs, or piloting proof-of-concept algorithms that attempt novel strategies.
+Further areas of research for general lossless data compression in the context of biological data could include exploring other properties like compression/decompression speed performance, meaning how fast data transcoding occurs, or exploring proof-of-concept algorithms that attempt novel strategies.
 
-# Supplementary information
+\newpage
 
-::: {.supp}
-## S1 Fig. {#s1-fig}
+# Supplementary Material
 
-Bold the title sentence.
+![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression.
+*Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation.
+In this instance, randomized DNA sequences were generated within a shorter range between 100 000 to 150 000 base pairs.
+The discrepancy between uncompressed and compressed data is less pronounced than in @fig-performance, though no compressed FASTA file exceeds more than four kilobytes in size.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotOne.png){#fig-rna}
 
-Add descriptive text after the title of the item (optional).
-:::
+![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression.
+*Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation.
+In this instance, randomized RNA sequences were generated within the range between 100 000 to 150 000 base pairs.
+The results visually remain similar to that of @fig-performance, although the compression performance of dictionary-based Zstd compression appears to have slightly regressed, perhaps due to some improvements in data compression gained via dictionary use no longer being possible due the difference in sequence type (substitution of thymine with uracil).](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotTwo.png){#fig-rna}
 
-::: {.supp}
-## S2 Fig. {#s2-fig}
+![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression.
+*Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation.
+In this instance, randomized amino acid sequences were generated within the range between 100 000 to 150 000 codons.
+Zstd default and dictionary-based compression performance remain similar, it being assumed that the more complex sequence type (20 possible codons versus four possible nucleotides per position) may have erased dictionary compression improvements trained from a DNA sequence dataset.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotThree.png){#fig-rna}
 
-Lorem ipsum.
-
-Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-:::
+\newpage
 
@@
 
-I would like to express my sincere gratitude to Professor David Walsh for teaching the bioinformatics course; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; Mark Kittisopikul for developing the CodecZstd.jl #dictionary_and_parameters branch for Zstd dictionary functions; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.
+I would like to express my sincere gratitude to Professor David Walsh for aiding me throughout the project and for teaching the bioinformatics course; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; Mark Kittisopikul for developing the CodecZstd.jl #dictionary_and_parameters branch for Zstd dictionary functions; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.
 

### 2022-12-10 paper/Persico2022.tex
Final paper, final code

--- 2022-12-10 paper/paper.tex
@@
 \section*{Abstract}
-\hypertarget{background}{%
-\subsection{Background}\label{background}}
-
 Zstandard (Zstd) represents a universal, lossless data compression
@@
 compression improvements and compare them to the compression performance
-of Xz, Zlib, Bzip2, and Lz4 universal compressors.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
+of Xz, Zlib, Bzip2, and Lz4 universal compressors. Zstd dictionary-based
+compression did not yield significant compression gains (P \textless{}
+0.05 in many compressor contrast scenarios using Tukey's HSD(honestly
+significant difference) test) and had performed worse than alternative
+compressor types. As an \emph{E.coli} training dataset was used to
+generate a dictionary to compress randomized data, and with the
+abundance of options available for manually tuning Zstd's parameters,
+further compression improvements may be possible.
 
@@
 
+\newpage
+
 \hypertarget{results}{%
@@
 
-\caption{\label{fig-performance}Average FASTA formatted data size
-(bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based
-Zstd compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
-artifact for Zstd dictionary generation. Randomized DNA sequences within
-the range of 100 000 to 150 000 base pairs.}
+\caption{\label{fig-performance}Average FASTA formatted data size, in
+bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
+compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
+artifact for Zstd dictionary generation{[}23{]}. Randomized DNA
+sequences within the range between 100 000 to 150 000 base pairs.}
 
 \end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Tukeytable.png}
+
+}
+
+\caption{\label{fig-tukey}Tukey's HSD test for determining statistical
+significance between the compression performance, based on compressed
+data size, of Bzip2, Xz, Zlib, Zstd with default parameters, and Zstd
+with a custom dictionary based on the E.coli training dataset{[}23{]}.}
+
+\end{figure}
+
+\newpage
+
+The Tukey's HSD test paints a picture of the statistical significance of
+the means of the compressed set of FASTA formatted data between each
+compressor type. As expected, uncompressed data size and compressed data
+size differ significantly, with no compressed FASTA file above more than
+60 kilobytes based on Fig~\ref{fig-performance}. With dictionary-based
+compression applied, Zstd performs similarly to Zstd compression with
+default parameters (P \textgreater{} 0.05) and, based on the trends of
+the boxplot and the Tukey's test results, in almost every instance aside
+from default Zstd compression, dictionary-based compression performed
+worse than the competing compressor types.
 
@@
 
-The provided FASTA training dataset was retrieved from a set of
+As stated in the Results section, Zstd dictionary-based compression
+results were surprising in the sense that, at least with the
+configuration (particular \emph{E.coli} dataset, specific number of
+randomized FASTA files, completely randomized DNA, RNA, or amino acid
+sequences), it performed either on par, as was the case with default
+Zstd compression, or worse than the alternative compressors. The
+provided FASTA training dataset was retrieved from a set of
 \textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome
@@
 compression/decompression performance when attempted on more dissimilar
-sequence data. Portability of Zstd dictionaries, in the sense of being
-universally applicable across all sequence data, could be weighed
-against generating dedicated dictionaries for specific datasets
-depending on the storage requirements. Zstd parameters were kept
-default, though they can be manually altered, including advanced
-compression options like compression job size or selecting from 22
-predefined compression levels\citep{facebook}.
+sequence data, which might explain the observed results as randomized
+sequences were generated for compression performance analysis.
+Portability of Zstd dictionaries, in the sense of being universally
+applicable across all sequence data, could be weighed against generating
+dedicated dictionaries for specific datasets depending on the storage
+requirements. All other Zstd parameters were kept default, though they
+can be manually altered, including advanced compression options like
+compression job size or selecting from 22 predefined compression
+levels\citep{facebook}.
 
@@
 implementation by altering available options (block size, compression
-level,\ldots) that might influence compression performance. The
+level,\ldots) might yield more significant compression performance. The
 development of novel data toolsets have utilized Zstd in unique ways.
@@
 other properties like compression/decompression speed performance,
-meaning how fast data transcoding occurs, or piloting proof-of-concept
+meaning how fast data transcoding occurs, or exploring proof-of-concept
 algorithms that attempt novel strategies.
 
-\hypertarget{supplementary-information}{%
-\section{Supplementary information}\label{supplementary-information}}
+\newpage
 
-\paragraph*{S1 Fig.}
-\label{s1-fig}
-{\textbf{Bold the title sentence.}} Add descriptive text after the title
-of the item (optional).
+\hypertarget{supplementary-material}{%
+\section{Supplementary Material}\label{supplementary-material}}
 
-\paragraph*{S2 Fig.}
-\label{s2-fig}
-{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
-gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
-ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
-consectetur pellentesque.
+\begin{figure}
+
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotOne.png}
+
+}
+
+\caption{\label{fig-rna}Average FASTA formatted data size, in bytes,
+following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
+compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
+artifact for Zstd dictionary generation. In this instance, randomized
+DNA sequences were generated within a shorter range between 100 000 to
+150 000 base pairs. The discrepancy between uncompressed and compressed
+data is less pronounced than in Fig~\ref{fig-performance}, though no
+compressed FASTA file exceeds more than four kilobytes in size.}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotTwo.png}
+
+}
+
+\caption{\label{fig-rna}Average FASTA formatted data size, in bytes,
+following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
+compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
+artifact for Zstd dictionary generation. In this instance, randomized
+RNA sequences were generated within the range between 100 000 to 150 000
+base pairs. The results visually remain similar to that of
+Fig~\ref{fig-performance}, although the compression performance of
+dictionary-based Zstd compression appears to have slightly regressed,
+perhaps due to some improvements in data compression gained via
+dictionary use no longer being possible due the difference in sequence
+type (substitution of thymine with uracil).}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotThree.png}
+
+}
+
+\caption{\label{fig-rna}Average FASTA formatted data size, in bytes,
+following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
+compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
+artifact for Zstd dictionary generation. In this instance, randomized
+amino acid sequences were generated within the range between 100 000 to
+150 000 codons. Zstd default and dictionary-based compression
+performance remain similar, it being assumed that the more complex
+sequence type (20 possible codons versus four possible nucleotides per
+position) may have erased dictionary compression improvements trained
+from a DNA sequence dataset.}
+
+\end{figure}
+
+\newpage
 
@@
 I would like to express my sincere gratitude to Professor David Walsh
-for teaching the bioinformatics course; My friends and family that
-supported me throughout my studies; The Julia community for their
-support before and throughout the writing of this paper; Mark
-Kittisopikul for developing the CodecZstd.jl
+for aiding me throughout the project and for teaching the bioinformatics
+course; My friends and family that supported me throughout my studies;
+The Julia community for their support before and throughout the writing
+of this paper; Mark Kittisopikul for developing the CodecZstd.jl
 \#dictionary\_and\_parameters branch for Zstd dictionary functions; The
